---
title: 'Chapter 2: Descriptive Statistics'
---

## 2.1 Stem-and-Leaf Graphs, Line Graphs, and Bar Graphs
One simple graph, the **stem-and-leaf graph** or **stemplot**, comes from the field of exploratory data analysis. It is a good choice when the data sets are small. To create the plot, divide each observation of data into a stem and a leaf. The leaf consists of a final **significant digit**. For example, 23 has stem two and leaf three. The number 432 has stem 43 and leaf two. Likewise, the number 5,432 has stem 543 and leaf two. The decimal 9.3 has stem nine and leaf three. Write the stems in a vertical line from smallest to largest. Draw a vertical line to the right of the stems. Then write the leaves in increasing order next to their corresponding stem.

**Try It: Problem 1**

For the Park City basketball team, scores for the last 30 games were as follows (smallest to largest): 

32; 32; 33; 34; 38; 40; 42; 
42; 43; 44; 46; 47; 47; 48; 
48; 48; 49; 50; 50; 51; 52; 
52; 52; 53; 54; 56; 57; 57; 
60; and 61.

Construct a stem plot for the data.

## Stem-and-Leaf Plot
```{r }
scores <- c(32, 32, 33, 34, 38, 40, 
            42, 42, 43, 44, 46, 47, 
            47, 48, 48, 48, 49, 50, 
            50, 51, 52, 52, 52, 53, 
            54, 56, 57, 57, 60, 61)

stem(scores)
```


**Try It: Problem 2**

The data are the distances (in kilometers) from a home to local supermarkets. Create a stemplot using the data: 

1.1; 1.5; 2.3; 2.5; 2.7; 
3.2; 3.3; 3.3; 3.5; 3.8; 
4.0; 4.2; 4.5; 4.5; 4.7; 
4.8; 5.5; 5.6; 6.5; 6.7; 12.3

Do the data seem to have any concentration of values?

## Stem-and-Leaf Plot
```{r }
scores <- c(1.1, 1.5, 2.3, 2.5, 2.7, 
            3.2, 3.3, 3.3, 3.5, 3.8, 
            4.0, 4.2, 4.5, 4.5, 4.7, 
            4.8, 5.5, 5.6, 6.5, 6.7, 12.3)

stem(scores, scale = 2)
```

Based on the data in the stemplot, the value **12.3** may be an outlier. Values appear to concentrate at three and four kilometers.


**Try It: Problem 3**

The following data show the distances (in miles) from the homes of off-campus statistics students to the college. Create a stem plot using the data and identify any outliers:

0.5; 0.7; 1.1; 1.2; 1.2; 
1.3; 1.3; 1.5; 1.5; 1.7; 
1.7; 1.8; 1.9; 2.0; 2.2; 
2.5; 2.6; 2.8; 2.8; 2.8; 
3.5; 3.8; 4.4; 4.8; 4.9; 
5.2; 5.5; 5.7; 5.8; 8.0

## Stem-and-Leaf Plot
```{r }
scores <- c(0.5, 0.7, 1.1, 1.2, 1.2, 
            1.3, 1.3, 1.5, 1.5, 1.7, 
            1.7, 1.8, 1.9, 2.0, 2.2, 
            2.5, 2.6, 2.8, 2.8, 2.8, 
            3.5, 3.8, 4.4, 4.8, 4.9, 
            5.2, 5.5, 5.7, 5.8, 8.0)

stem(scores, scale = 2)
```

The value **8.0** may be an outlier. Values appear to concentrate at one and two miles.